---
title: "Class 05: Data Visualization with GGPLOT"
author: "Zixuan Zeng (PID: 16142927)"
format: pdf
---

Today we r playing with plotting and graphics in R.

There are lots of ways to make cool figures in R.
There is "base" R graphics (`plot()`,`hist()`,`boxplot()` etc.)

There is also add-on packages, like **ggplot**

```{r}
head(cars)
```
Let's plot this with "base" R:

```{r}
head(mtcars)
```
Let's plot `mpg` vs `disp`

```{r}
plot(mtcars$mpg,mtcars$disp, pch=16)
```

```{r}
hist(mtcars$mpg)
```

##GGPLOT

The main function in the ggplot2 package is
**ggplot()**
`install.packages()`
The ggplot2 package was installed before this lab session.
>**N.B.** never run `install.packages()` in quarto source document



```{r}
library(ggplot2)
ggplot(cars) + aes(speed,dist) + geom_point()
```
```{r}
ggplot(cars) + aes(speed) + geom_histogram()
```

Every ggplot needs at least 3 things:

-The **data**
-The **aes** mapping
-The **geom**
>For simple canned graphs "base" R is nearly always faster

Add a line, a tile, a subtitle, and caption as well as custom axis labels

```{r}
ggplot(cars) + aes(speed,dist,) + geom_point() + geom_smooth(method = "lm",se=FALSE) + labs(title="silly plot", x = "Speed (MPH)", y = "Distance(ft)", caption = "when is tea time anyway?") + theme_bw()
```

## plot some expresion data

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

> Q1. How many genes are in this wee dataset?

There are `r nrow(genes)` in this dataset.

> Q2. How many "up" regulated genes are there?

```{r}
sum(genes$State == "up")
```

```{r}
table(genes$State)
```

```{r}
p <- ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point() +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red"))
```


```{r}
p + labs(title = "look at me")
```

Silly example of adding labels

```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State, label=Gene) + geom_point() +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red")) + geom_text()
```

```{r}
library(ggrepel)
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State, label=Gene) + geom_point() +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red")) + geom_text_repel(max.overlaps = 60) + theme_bw()
```


##going further

Playing with some different layers and the gapmider dataset...

```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

```{r}
head(gapminder)
```


```{r}
tail(gapminder)
```

```{r}
ggplot(gapminder) + aes(y=lifeExp,x=gdpPercap, col=continent) + geom_point() + facet_wrap(~continent) + theme_bw()
```

```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point() +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red")) + facet_grid(~State)
```

```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_point(alpha = 0.5) +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red")) 
```

```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_violin() +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red"))
```


```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_bin_2d() +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red"))
```

```{r}
ggplot(genes) + aes(x=Condition1, y=Condition2, col=State) + geom_rug() +labs(x="Control",y="drug treatment") + scale_color_manual(values=c("blue","gray","red"))
```

