---
title: "Class6:R functions"
author: "Zixuan Zeng (A16142927)"
format: pdf
toc: true
---

All functions in R have at least 3 things:

- A **name**, we oucj this and use it to call our function,
-Input **arguments** (there can be multiple)
The **body** lines o R code that do the work


## Our first (silly) function

write a function to add some numbers

```{r}
add <- function(x,y=1) {
  x + y
}
```

Now we can call the function:
```{r}
add(c(10,10),100)
```


```{r}
add(10,100)
```
## A second function

Write a function to generate random nucleotide sequences of a user specified length:

THe `sample()` function can be helpful here.

```{r}
V <- sample(c("A","C","G","T"),size=50, replace = TRUE)
```
I want the a 1 element long character vector that looks "TCATTG" not "T" "C" "A" "T" "T" "G"

```{r}
V <- sample(c("A","C","G","T"),size=50, replace = TRUE)
paste(V,collapse= "")
```
Turn this into generate_DNA
```{r}
generate_DNA <- function(size = 50){
V <- sample(c("A","C","G","T"),size=size, replace = TRUE)
paste(V,collapse= "")
}
```

Test it:
```{r}
generate_DNA(60)
```

```{r}
fasta <- FALSE
if(fasta){
  cat("HELLO You!")
} else{
  cat("No you dont!")
}
```

Add the ability to return a multi-element vector or a single element fasta like vector.

```{r}
generate_DNA <- function(size = 50, fasta=TRUE){
V <- sample(c("A","C","G","T"),size=size, replace = TRUE)
s <- paste(V,collapse= "")

if(fasta){
  return(s)
} else{
  return(V)
}
}
```

Test:
```{r}
generate_DNA(60,fasta=FALSE)
```

## a protein generating function

```{r}
generate_protein <- function(size = 50, fasta=TRUE){
V <- sample(c("A", "R", "N", "D", "C", "E", "Q", "G", "H", "I", "L", "K", "M", "F", "P", "S", "T", "W", "Y", "V"),size=size, replace = TRUE)
s <- paste(V,collapse= "")

if(fasta){
  return(s)
} else{
  return(V)
}
}
generate_protein(6)
```
  
  
Use our new `generate_protein`function to make random protein sequences of length 6 to 12 (i.e. one length 6, one length 7, up to length 12)

brute force:
```{r}
generate_protein(6)
generate_protein(7)
generate_protein(8)
```

A second way is to use a `for()` loop:
```{r}
lengths <- 6:12
lengths
```

```{r}
for(i in lengths){
  cat(">",i,"\n",sep="")
  aa <- generate_protein(i)
  cat(aa)
  cat("\n")
}
```
A thrid, and better, way to solve this is to use the `apply()` family of functions,specifically the `sapply()` function in this case
```{r}
sapply(6:12,generate_protein)
```


  
  